---
title: "Inference for numerical data"
output: statsr:::statswithr_lab
---

Here are my answers to the Week 3 Lab Activity of the couse __Inferential Statistics with R__ presented by Coursera and conducted by Mine Çetinkaya-Rundel.

## R Markdown

This is a R Markdown file. To a better viewing, it could be forked and knitted on [RStudio](https://www.rstudio.com/) to a html file or it could be viewed directly as a [RPubs publication](http://rpubs.com/rodsveiga/252171). 

## Packages

We will use the `devtools` package to install the `statsr` package associated with this course. We need to install and load this package.

```{r, eval=FALSE}
install.packages("devtools")
library(devtools)

```

Now we can install the rest of the packages we will use during the course. Type the following commands in the Console as well:

```{r, eval=FALSE}
install.packages("dplyr")
install.packages("ggplot2")
install.packages("shiny")
install_github("StatsWithR/statsr")

```

## Getting Started

### Load packages

In this lab we will explore the data using the `dplyr` package and visualize it 
using the `ggplot2` package for data visualization. The data can be found in the
companion package for this course, `statsr`.

Let's load the packages.

```{r load-packages, message=FALSE}
library(statsr)
library(dplyr)
library(ggplot2)
```

### The data

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

Load the `nc` data set into our workspace.

```{r load-data}
data(nc)
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | ---------------------------------------------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

```{r}
names(nc)
head(nc)
```

Note there are some missing data for the variable `fage`, indicanting there are single mothers in this data set.

1. There are 1,000 cases in this data set, what do the cases represent? 
<ol>
<li> The hospitals where the births took place </li> 
<li> The fathers of the children </li> 
<li> The days of the births </li>
<li> The births </li>
These 1,000 cases is the data set represent the births.
</ol>

As a first step in the analysis, we should take a look at the variables in the dataset. 
This can be done using the `str` command:

```{r str}
str(nc)
```

As you review the variable summaries, consider which variables are categorical and which 
are numerical. For numerical variables, are there outliers? If you aren't sure or want to 
take a closer look at the data, make a graph.

The `Factor` variables are categoricalm while the `int` and `num` are numerical.

## Exploratory data analysis

We will first start with analyzing the weight gained by mothers throughout the 
pregnancy: `gained`.

Using visualization and summary statistics, describe the distribution of weight 
gained by mothers during pregnancy. The `summary` function can also be useful.

```{r summary}
summary(nc$gained)
```

2. How many mothers are we missing weight gain data from?
<ol>
<li> 0 </li>
<li> 13 </li>
<li> 27 </li>
<li> 31 </li>
The are 27 missing values in the weight gain data.
</ol>

Next, consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

3. Make side-by-side boxplots of `habit` and `weight`. Which of the following is 
false about the relationship between habit and weight?
<ol>
<li> Median birth weight of babies born to non-smoker mothers is slightly higher than that of babies born to smoker mothers. </li>
<li> Range of birth weights of babies born to non-smoker mothers is greater than that of babies born to smoker mothers. </li>
<li> Both distributions are extremely right skewed. </li>
<li> The IQRs of the distributions are roughly equal. </li>
Only the nonsmoker distribution is extremely right skewed. There are three outliers in the smoker distribution, but as the sample size is large we should not concern about extremely skew.  
</ol>
```{r habit-weight-box}
# nc$weight is a numerical vector of data values to be split into groups according to the grouping variable nc$habit (factor).
boxplot(weight ~ habit, data = nc, xlab = "Habit", ylab = "Weight")
# We can compare the two groups using the by() function together with the summary() function.
by(nc$weight, nc$habit, summary)
# We could be interested in the stardard deviations and the IQR's as well.
by(nc$weight, nc$habit, sd)
# As expected by checking the summary, the IQRs of the distributions are roughly equal.
by(nc$weight, nc$habit, IQR)
```
```{r}

```

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following to 
first group the data by the `habit` variable, and then calculate the mean
`weight` in these groups using the `mean` function.

```{r by-means}
nc %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight))
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis 
test.

```{r ttest}
# We can make a T test using the R funcrion t.test().
t.test(weight ~ habit, data = nc, conf.level = 0.95)
```

Below we are goint to be introduced to the `inference` function. One could check that its results for the confidence interval are very close to those of `t.test`, but they are not exactly the same. One can also note that the degrees of freedom used in both calculations are different. This gives us a clue that this slightly deviation comes from their internal method of calculation.  While the `t.test` documentation does not give information about how the calculation takes place, the `inference` function has a `method` option, which can be chosen as `theoretical` $(df = min(n_1 -1, n_2 -2))$ or `simulation` (randomization/bootstrap).



## Inference


<div id="exercise">
**Exercise**: Are all conditions necessary for inference satisfied? Comment on each. You can compute the group sizes using the same `by` command above but replacing `mean(weight)` with `n()`.
</div>

We have no reason to believe the sample observations within the groups are not independent, as well the samples groups. Certainly, the samples sizes are less than 10% of the population size. 

The samples size are large enough, so we do not have to worry about skew.

```{r}
# Computing the group size.
nc %>%
   group_by(habit) %>%
   summarise(group_size = n())
```


4.  What are the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different?
<ol>
<li> $H_0: \mu_{smoking} = \mu_{non-smoking}$; $H_A: \mu_{smoking} > \mu_{non-smoking}$ </li>
<li> $H_0: \mu_{smoking} = \mu_{non-smoking}$; $H_A: \mu_{smoking} \ne \mu_{non-smoking}$ </li>
<li> $H_0: \bar{x}_{smoking} = \bar{x}_{non-smoking}$; $H_A: \bar{x}_{smoking} > \bar{x}_{non-smoking}$ </li>
<li> $H_0: \bar{x}_{smoking} = \bar{x}_{non-smoking}$; $H_A: \bar{x}_{smoking} > \bar{x}_{non-smoking}$ </li>
<li> $H_0: \mu_{smoking} \ne \mu_{non-smoking}$;  $H_A: \mu_{smoking} = \mu_{non-smoking}$ </li>
The hypotheses test is given by $H_0: \mu_{smoking} = \mu_{non-smoking}$; $H_A: \mu_{smoking} \ne \mu_{non-smoking}$.
</ol>

Next, we introduce a new function, `inference`, that we will use for conducting
hypothesis tests and constructing confidence intervals. 

Then, run the following:

```{r inf-weight-habit-ht, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `habit`. The third argument, `data`, is the data frame these
variables are stored in. Next is `statistic`, which is the sample statistic
we're using, or similarly, the population parameter we're estimating. In future labs
we can also work with "median" and "proportion". Next we decide on the `type` of inference 
we want: a hypothesis test (`"ht"`) or a confidence interval (`"ci"`). When performing a 
hypothesis test, we also need to supply the `null` value, which in this case is `0`, 
since the null hypothesis sets the two population means equal to each other. 
The `alternative` hypothesis can be `"less"`, `"greater"`, or `"twosided"`. 
Lastly, the `method` of inference can be `"theoretical"` or `"simulation"` based.

For more information on the inference function see the help file with `?inference`.

<div id="exercise">
**Exercise**: What is the conclusion of the hypothesis test?
</div>

As the p-value is smaller than the usual confidence level 0.05, we reject the null hypotheses. This means this data set provides enought evidence that the average weights of babies born to smoking and non-smoking mothers are different.

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to nonsmoking and 
smoking mothers, and interpret this interval in context of the data. Note that by 
default you'll get a 95% confidence interval. If you want to change the
confidence level, add a new argument (`conf_level`) which takes on a value
between 0 and 1. Also note that when doing a confidence interval arguments like
`null` and `alternative` are not useful, so make sure to remove them.
<ol>
<li> We are 95% confident that babies born to nonsmoker mothers are on average 0.05 to 0.58 pounds lighter at birth than babies born to smoker mothers. </li>
<li> We are 95% confident that the difference in average weights of babies whose moms are smokers and nonsmokers is between 0.05 to 0.58 pounds. </li>
<li> We are 95% confident that the difference in average weights of babies in this sample whose moms are smokers and nonsmokers is between 0.05 to 0.58 pounds. </li>
<li> We are 95% confident that babies born to nonsmoker mothers are on average 0.05 to 0.58 pounds heavier at birth than babies born to smoker mothers. </li>
Taking into account the definition of confidence interval and running the R code below, we can say we are 95% confident that babies born to nonsmoker mothers are on average 0.05 to 0.58 pounds heavier at birth than babies born to smoker mothers.
</ol>
```{r habit-weight-ci}
# Te response variable is named as y and the exploratory variable as x.
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ci", conf_level = 0.95 ,method = "theoretical")
```

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$)
. We can easily change this order by using the `order` argument:

```{r inf-weight-habit-ci, tidy=FALSE}
inference(y = weight, x = habit, data = nc, statistic = "mean", type = "ci", 
          method = "theoretical", order = c("smoker","nonsmoker"))
```

6. Calculate a 99% confidence interval for the average length of pregnancies 
(`weeks`). Note that since you're doing inference on a single population 
parameter, there is no explanatory variable, so you can omit the `x` variable 
from the function. Which of the following is the correct interpretation of this 
interval?
<ol>
<li>(38.1526 , 38.5168)</li>
<li>(38.0892 , 38.5661)</li>
<li>(6.9779 , 7.2241)</li>
<li>(38.0952 , 38.5742)</li>
Running the code below we find that the correct answer for 99% confidence interval for the average lenght of prenancies is (38.0952 , 38.5742).
</ol>
```{r weeks-ci-99}
inference(y = weeks, data = nc, statistic = "mean", type = "ci", conf_level = 0.99 ,method = "theoretical")
```

<div id="exercise">
**Exercise**: Calculate a new confidence interval for the same parameter at the 90% confidence level. Comment on the width of this interval versus the one obtained in the the previous exercise.
</div>
```{r weeks-ci-90}
inference(y = weeks, data = nc, statistic = "mean", type = "ci", conf_level = 0.9 ,method = "theoretical")
```

As expected, the 90% confidence interval is narrower than the 95% confidence interval, since its confidence level is lower.

<div id="exercise">
**Exercise**: Conduct a hypothesis test evaluating whether the average weight gained by younger mothers is different than the average weight gained by mature mothers.
</div>
```{r gained-mature-ht}
inference(y = gained, x = mature, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")

```

As the p-value is greater than the significance level, we fail to reject the null hypothesis (the average weight gained by younger mothers is equal to the average weight gained by mature mothers).

We could have used the `t.test` fucntion:

```{r}
t.test(gained ~ mature, data = nc, conf.level = 0.95)
```


7. Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.
```{r cutoff-mature}
by(nc$mage, nc$mature, summary)
```

Note that the maximum age of younger moms is 34 and minimum age of mature moms is 35, which suggests us that the age cutoff for younger moms is 35. 

```{r final-boxplot}
boxplot(mage ~ mature, data = nc, ylab = "Mother's age in years")
```


<div id="exercise">
**Exercise**: Pick a pair of variables: one numerical (response) and one categorical 
(explanatory). Come up with a research question evaluating the relationship between 
these variables. Formulate the question in a way that it can be answered using a 
hypothesis test and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language. Be sure to check all assumptions,state your $\alpha$ level, and conclude 
in context. (Note: Picking your own variables, coming up with a research question,
and analyzing the data to answer this question is basically what you'll need to do for
your project as well.)
</div>

We choose to work with the variables `visits` (response) and `premie` (explanatory). We want to know whether the data set provides statistical significant evidence that the number of hospital visits is, on average, greater for the full term births than the premature births.  

```{r pick-your-own}
boxplot(visits ~ premie, data = nc, xlab = "Birth classification", ylab = "Hospital visits during pregnancy")

```

We will conduct the following hypothesis test:

\[  H_0 : \ \   \mu_{full term} =  \mu_{premie}  \]  
\[  H_A : \ \  \mu_{full term} > \mu_{premie}   \] 

```{r pick-your-own2}
inference(y = visits, x = premie, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "greater", method = "theoretical")
```

As the p-value is really small, we reject the null hypothesis (we suppose that a reasonable confidence level is greater than 0.0001). This means the data set provides statistical significant evidence that the number of hospital visits is, on average, greater for the full term births than the premature births.

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>